Functional pant indicators - wetland, indicator documentation

### 2023-04-21 DRAFT_functional plant indicators_wetland.Rmd
finishing up first fraft of the wetland file

<!-- Load all you dependencies here -->
```{r setup, include=FALSE}
library(sf)
library(tidyr)
library(dplyr)
library(plyr)
library(stringr)
library(tidyverse)
library(readxl)
library(broom)
library(RColorBrewer)
library("gridExtra") 
library(ggridges)
library(ggplot2)
library(tmap)
library(tmaptools)
library(raster)
library(stars)

@@

We then can prepare a list of data frames to hold the results and perform the scaling according to the principles described in NINA report 1967 (Töpper and Jakobsson 2021)
```{r, echo = T, eval = F}

@@

# roll out
results.wetland[['scaled']] <- results.wetland[['non-truncated']] <- results.wetland[['original']]

@@

We can calculate simple summary statistics like means, standard deviations, and number of observations (note that there is both a lower and an upper indicator for each of the four plant functional indicators)

@@

And we can show the resulting scaled values as Violin plots for each indicator and main ecosystem type 
```{r}

@@

We can also show the results as a map, for instance for pH1 (the lower pH indicator), either by directly plotting the data onto the map...
```{r, message = FALSE}
#### scaled value maps ####
# keep wide format and add geometry again
res.wetland2 <- results.wetland[['2-sided']]
st_geometry(res.wetland2) <- st_geometry(ANO.wetland)

nor <- readRDS('P:/41201785_okologisk_tilstand_2022_2023/data/rds/norway_outline.RDS')%>%
  st_as_sf() %>%
  st_transform(crs = crs(ANO.geo))
reg <- st_read("P:/41201785_okologisk_tilstand_2022_2023/data/regioner/regNorway_wgs84 - MERGED.shp")%>%
  st_as_sf() %>%
  st_transform(crs = crs(ANO.geo))
# change region names to something R-friendly
reg$region
reg$region <- c("Northern Norway","Central Norway","Eastern Norway","Western Norway","Southern Norway")

@@

```
...but here the colors and values of the data points on the map are hard to make out.
Alternatively we can calculate and show the region-wise means and standard deviations
```{r}

@@

#### CONTINUE HERE ####

#### Regions
<!-- In case you need to map the indicator value to regions, you can do that here. Remove this chapter if not relevant. -->
Importing a shape file with the regional delineation. 
```{r}
#reg <- sf::st_read("data/regions.shp")
#st_crs(reg)
```

### Scaled indicator values
<!-- Text and analyses here -->

### Uncertainty
<!-- Text here -->

## Prepare export
<!-- Text here -->

>>>>

<!-- Load all you dependencies here -->
```{r setup, include=FALSE}
library(sf)
library(tidyr)
library(plyr)
library(dplyr)
library(stringr)
library(tidyverse)
library(readxl)
library(broom)
library(RColorBrewer)
library("gridExtra") 
library(ggridges)
library(ggplot2)
library(tmap)
library(tmaptools)
library(raster)
library(stars)

@@

We then can prepare a list of data frames to hold the results and perform the scaling according to the principles described in NINA report 1967 (Töpper and Jakobsson 2021)
```{r, include=FALSE}

@@

# roll out
results.wetland[['scaled']] <- results.wetland[['non-truncated']] <- results.wetland[['original']]
```

```{r, echo = T, eval = F}

@@

### Simple summary statistics
We can calculate simple summary statistics like means, standard deviations, and number of observations (note that there is both a lower and an upper indicator for each of the four plant functional indicators)

@@

### Ecosystem sub-types
And we can show the resulting scaled values as Violin plots for each indicator and main ecosystem type 
```{r, message=FALSE}

@@

### Indicator index maps
We can also show the results as a map, for instance for pH1 (the lower pH indicator), either by directly plotting the data onto the map...
```{r, message = FALSE}
#### scaled value maps ####
# keep wide format and add geometry again
res.wetland2 <- results.wetland[['2-sided']]
st_geometry(res.wetland2) <- st_geometry(ANO.wetland)

#nor <- readRDS('P:/41201785_okologisk_tilstand_2022_2023/data/rds/norway_outline.RDS')%>%
#  st_as_sf() %>%
#  st_transform(crs = crs(ANO.geo))

nor <- st_read("data/outlineOfNorway_EPSG25833.shp")%>%
  st_as_sf() %>%
  st_transform(crs = crs(ANO.geo))

#reg <- st_read("P:/41201785_okologisk_tilstand_2022_2023/data/regioner/regNorway_wgs84 - MERGED.shp")%>%
#  st_as_sf() %>%
#  st_transform(crs = crs(ANO.geo))

reg <- st_read("data/regions.shp")%>%
  st_as_sf() %>%
  st_transform(crs = crs(ANO.geo))

# change region names to something R-friendly
reg$region
reg$region <- c("Northern Norway","Central Norway","Eastern Norway","Western Norway","Southern Norway")

@@

```
...but here the colors and values of the data points on the map are hard to make out.

### Regions
Alternatively we can calculate and show the region-wise means and standard deviations
```{r}

@@
